---
title: "Categorical Variables"

output: 
    rmarkdown::html_vignette:
      toc: true
vignette: >
  %\VignetteIndexEntry{categorical}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}

---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#ws>"
)
```

```{r setup}
library(Intro2MLR)
```


# Introduction

The use of categorical variables is an important part of MLR and regression in general. Therefore we must learn about them and the multinomial.

## Example 1 MS pg 448 9.8

### Prepare the data as a table

```{r}
data(irr)
irr
irr$FREQ -> opinion
names(opinion) <- irr$STRATEGY
opinion <- as.table(opinion)
opinion
```

### Now analyze

Suppose we wish to test the NULL hypothesis

$$H_0:p_i = 1/5,\;\;\forall i\in\{1,\ldots,5\}$$

```{r}
test <-chisq.test(opinion)
test
```

## Conclusion

We do not have sufficient evidence at the 0.05 level to reject the NULL hypothesis and so we retain it as plausibly true given the data.

## Going further

You can investigate the $\chi^2$ test and associated statistics further by looking at the output

```{r}
names(test)
```
Notice that there are some very useful components that make up the test.

### The expected values $E(n_i)$

```{r}
test$expected
```

These are calculated under the assumption that the NULL hypothesis is correct
