---
title: "The F statistic in regression"

output: 
    rmarkdown::html_vignette:
      toc: true
vignette: >
  %\VignetteIndexEntry{F-stat}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}

---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#ws>"
)
```

```{r setup}
library(Intro2MLR)
```


# Introduction

The F statistic is central to MLR and we will need to both derive and use it well.

# Expressions for F

The test for model adequacy is usually presented as the test against 

$$H_0: \beta_1=\beta_2=\ldots =\beta_k$$
The F statistic is calculated by finding


$$F = \frac{(RSS_{H_0}-RSS)/k}{RSS/(n-(k+1))}$$
Now $RSS_{H_0}=(Y-\hat{Y})^{'}(Y-\hat{Y})$ where $\hat{Y}$ is calculated under the assumption of $H_0$.

So $Y=X\beta + \epsilon$ where $X= 1_n$ and hence $\hat{Y}=X\hat{\beta}$

$$\hat{Y}=X(X^{'}X)^{-1}X^{'} Y = 1_{n}(n)^{-1}1^{'}_n Y=1/nJ_n Y=\bar{Y}$$

This means that $RSS_{H_0}=TSS$

Since $TSS=MSS + RSS$ 

$$F = \frac{MSS/k}{RSS/(n-(k+1))}$$
