---
title: "FLAG"
output:
  html_document:
    toc: yes
    df_print: paged
  rmarkdown::html_vignette:
    toc: yes
vignette: |
  %\VignetteIndexEntry{FLAG} %\VignetteEngine{knitr::rmarkdown} %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center",fig.width = 6)
```

# Introduction

On page 627 of MS we have the account of the cost of a road construction awarded using the sealed bid system.


# Multi-colinearity

We will first investigate the predictor variables.

We will remove the `STATUS` variable since it is qualitative and we wish to assess quantitative correlation using `cor`.

```{r}
library(Intro2MLR)
library(dplyr)
data(flag)
nms <- names(flag)

cflag <- cor(flag)

cflag

```
This is helpful in many ways -- we note that some of the quantitative variables are highly correlated and that the response `COST` is correlated to a number of predictors.

```{r}
as.data.frame(cflag) %>% filter(COST>0.7 & COST < 1)
```



Even better we can make a correlation plot

```{r}
library(corrplot)
corrplot(cflag)
```



We will remove `DAYSEST` and retain `DOTEST`. We need to investigate the relationship of the response to the qualitative variable `STATUS`

```{r}
library(ggplot2)
flag <- flag %>% mutate(STATUS <- factor(STATUS))
g <- ggplot(flag, aes(x = STATUS, y = COST)) + geom_point(aes(col = COST)) 
g

```

The above plot shows that the mean cost changes as the levels (0 -> 1) change. We will therefore include the `STATUS` qualitative variable.

# Model

We will trial the following model

$$ COST \sim DOTEST + STATUS + STATUS:DOTEST $$

# In R

```{r}
ylm <- lm(COST ~ DOTEST + STATUS + STATUS:DOTEST, data = flag)
summary(ylm)
```


We will need to investigate the validity of the model before progressing - the model as it is shows adequacy.

## Validity

### Normality

Check on normality $H_0: \epsilon \sim N$

```{r}
library(s20x)
normcheck(ylm, shapiro.wilk = TRUE)
```

The residuals give evidence against normality.

### Constant variance

```{r}
plot(ylm, which = 1)
```


This is sad! Notice the fan!!!

We will try transforming the response.


```{r}


ylm1 <- lm(log(COST) ~ DOTEST + STATUS + STATUS:DOTEST, data = flag)
plot(ylm1, which = 1)

```

This worked on the fan but now we have a trend in the residuals.

We will try transforming the predictor 


```{r}

ylm2 <- lm(log(COST) ~ log(DOTEST) + STATUS + STATUS:I(log(DOTEST)), data = flag)
plot(ylm2, which = 1)
```

As can be seen by the plot of residuals versus fitted values there is no trend and points are evenly scattered about the horizontal axis.

Cool!!

## Simplify model

We may find that the model needs to be simplified now.

```{r}
summary(ylm2)
```

Sure enough we can remove the interaction term


```{r}
ylm3 <- lm(log(COST) ~ log(DOTEST) + STATUS , data = flag)
```



Now we will recheck normality

```{r}
normcheck(ylm3, shapiro.wilk = TRUE)
```

The QQ plot and Normal plot alongside are better. We still have evidence against the Normal assumption but the model has improved.

# Prediction

Because we have transformed the response we will need to back transform our predictions to get answers on the original scale.

```{r}
summary(ylm3)
```

From the output we can write:

$$\widehat{log(COST)} = -0.147 + 1.01 log(DOTEST) + 0.217 STATUS$$
Suppose we wish to predict the road construction cost when the DOT estimate is \$370,000 this means `DOTEST=370` and contract is fixed `STATUS=1`

Then you can use the above equation or use `predict`

```{r}
ans = predict(ylm3, data.frame(DOTEST=370, STATUS = 1), interval = "none")
ans
```


This is on the log scale so on the original scale we will have

```{r}
exp(ans)
```

So the cost for a fixed road contract with DOT estimate of \$370,000 will be \$`r exp(ans)` thousand dollars.

This is not that satisfying statistically however, since a point estimate without and interval is not even half the story.

# Interval estimate

The prediction interval for a new response is given below with the transformed 95\% interval.

```{r}
ans1 = predict(ylm3, data.frame(DOTEST=370, STATUS = 1), interval = "prediction")
ans1
exp(ans1)
```

# ANOVA

You should understand the subdivision of the TSS in terms of component RSS differences in sub models.

```{r}
anova(ylm3)
r<-residuals(ylm3)
sum(r^2)
```

