---
title: "Using Flexdashboard"
output: 
  rmarkdown::html_vignette:
    toc: true
    
vignette: >
  %\VignetteIndexEntry{Flexdashboard}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#ws>"
)
```

```{r setup}
library(Intro2MLR)
```


# Introduction

We will often wish to present our findings at conferences and other gatherings where a good presentation of tables and graphics are required.

Flexdashboard is a great way to accomplish this.

To get up to speed in this area you will need to make good use of the online resources.

<https://rmarkdown.rstudio.com/flexdashboard/>


I will outline the basic ideas and you can perfect the details by inspecting the above website.


The first thing to notice is that there is an orientation

## Orientation

You can either orient the document by columns (default) or rows. This can be set in the `YAML` header using the keyword `orientation` which you can set as `rows` or `columns`.

### Example

```
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    
```

This will cause that all charts will populate the knitted document row-wise.

Whereas the following

```
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    
```
will populate the knitted document by columns.

## The meaning of headings

Normally when you make a document you would have `# Heading ` then `## Sub heading ` and then `### Sub Sub heading`

In `flex_dashboard` the hashes mean something different.

## Hash headings

The first hash `# page` denotes a new page (this will appear in a menu)

The second hash `## Column` will produce a new column if the orientation was columns. This can be denoted differently. You could use 
```
Column
------------------------------------
```

The third hash `### Chart` would make space for a chart or table etc


# Example 

what would the following dashboard code produce when knitted?

```
---
title: "Column Orientation"
output: flexdashboard::flex_dashboard
---

    
Column
-------------------------------------
    
### Chart 1
    
```{r}
```
   
Column
-------------------------------------
   
### Chart 2

```{r}
```   
 
### Chart 3
    
```{r}
```

```

Notice that the YAML header has no orientation set. So it will be the default orientation: `columns`

The second order header:

```
Column
-------------------------------------
    
```
Will prepare a column for the first chart which is set with the three hashes

```
### Chart 1

```{r}
```   
 
```
The last code produces a new column into which two charts 2, 3 will be rendered.

```
Column
-------------------------------------
   
### Chart 2

```{r}
```   
 
### Chart 3
    
```{r}
```

```

# Problem 1

We will use our current understanding to solve a problem.

Suppose we want want to have a row of 2 charts followed by one row containing a third chart.

Produce the code to make the document.


```
---
title: "Problem 1"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
---
    
Row
-------------------------------------
    
### Chart 1
    
```{r}
```
   
   
### Chart 2  

Row
-------------------------------------
    
### Chart 3
    
```{r}
```
    
```


# Problem 2

Suppose we wish to have a large number of charts on a page. This can be accomplished by using tabsets. You probably are familiar with  these in your browsers.

Produce the code that will create three tabset charts in one row and 4 tabset charts in a second row.


```
---
title: "Tabset Column"
output: flexdashboard::flex_dashboard:
          orientation: rows
---
    
Column {.tabset}
-------------------------------------
    
### Chart 1
    
```{r}
```

### Chart 2
    
```{r}
```

### Chart 3
    
```{r}
```


   
Column {.tabset}
-------------------------------------
   
### Chart 4

```{r}
```   
 
### Chart 5
    
```{r}
```


### Chart 6
    
```{r}
```

```



