---
title: "Example of multicollinearity"

output: 
    rmarkdown::html_vignette:
      toc: true
vignette: >
  %\VignetteIndexEntry{multicol}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Multi-collinearity can be found in the following regression.

# FTC2 data

```{r,cache=TRUE}
v<-Intro2MLR::myreadxl()
ftc2<-v$FTC2
rm(v)
```



## Model 1

```{r}
names(ftc2)
fit1 <- lm(CO ~ TAR + NICOTINE + WEIGHT, data = ftc2)
summary(fit1)
car::vif(fit1)
cor(ftc2[,1:3])
```

## New model

We should remove the X with most variance inflation.  Note that NICOTINE is highly correlated with TAR. So we should remove NICOTINE or TAR -- we will choose to remove NICOTINE since it has the largest VIF.


```{r}
fit2 <- lm(CO ~ TAR + WEIGHT, data = ftc2)
summary(fit2)
car::vif(fit2)

```

## Remove WEIGHT

Weight is shows very little evidence that it will impact predictions (see P value and T stats).

We will remove WEIGHT



```{r}
fit3 <- lm(CO ~ TAR , data = ftc2)
summary(fit3)
anova(fit3,fit2)
```
## $R^2_a$

We will now trace the adjusted R squared.

```{r}
sm1<-summary(fit1)
sm2<-summary(fit2)
sm3<-summary(fit3)
sm1$adj.r.squared
sm2$adj.r.squared
sm3$adj.r.squared
```

The last model has the largest adjusted R squared.

