---
title: "multicol"

output: 
    rmarkdown::html_vignette:
      toc: true
vignette: >
  %\VignetteIndexEntry{multicol}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Multi-collinearity can be found in the following regression.

# FTC2 data

```{r}
library(Intro2MLR)
data(ftc2)
```



## Model 1

```{r}
names(ftc2)
fit1 <- lm(CO ~ TAR + NICOTINE + WEIGHT, data = ftc2)
summary(fit1)
car::vif(fit1)
cor(ftc2[,1:3])
```

## New model

We should remove the X with most variance inflation.  Note that NICOTINE is highly correlated with TAR. So we should remove NICOTINE or TAR -- we will choose to remove NICOTINE since it has the largest VIF. We will use the `car` package calculate the VIF

$$VIF_i = \frac{1}{1-R^2_i}$$


```{r}
fit2 <- lm(CO ~ TAR + WEIGHT, data = ftc2)
summary(fit2)
car::vif(fit2)

```

## Remove WEIGHT

Weight is shows very little evidence that it will impact predictions (see P value and T stats).

We will remove WEIGHT



```{r}
fit3 <- lm(CO ~ TAR , data = ftc2)
summary(fit3)
anova(fit3,fit2)
```

Notice that the test produced by the `anova` function produces evidence against:

$$H_0: \beta_2 = 0$$
where $\beta_2$ is the coefficient of `WEIGHT`. You need to see that this is a situation where `k-g=1` and hence the F test corresponds to a two tailed T test, where $F=T^2$.

```{r}
sm1<-summary(fit1)
sm2<-summary(fit2)
sm3<-summary(fit3)

sm2$coefficients[3,]
```


In this case $f_{calc} = t_{calc}^2 = `r sm2$coefficients[3,3]^2`$ and the two pvalues are identical.

## $R^2_a$

We will now trace the adjusted R squared.

```{r}

sm1$adj.r.squared
sm2$adj.r.squared
sm3$adj.r.squared
```

The last model has the largest adjusted R squared.

